---
title: "Distributional Consequences of Climate Policy- ESM 204 Assignment 3"
author: "Sam Rozal, Hollie Pennington, Abigail Sanford"
date: "May 12th, 2022"
output:
  html_document:
    code_folding: hide
---

```{r setup, include = TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(rootSolve)
library(janitor)
library(here)
```

```{r}
df <- read_csv(here("data","HW3_data.csv")) %>% 
  select(-1) %>% 
  clean_names()
```
Last year, the Biden Administration assembled an Inter-agency Working Group (IWG) tasked with updating the United States Government’s Social Cost of Carbon (SCC), which has not been comprehensively updated since 2010. The Administration has also called on government agencies to address environmental justice, racism, and equity concerns when considering policies designed to mitigate climate change.

While the Interagency Working Group develops a new SCC, the Biden Administration’s “interim” value is $51 per metric ton of CO2. The electricity sector is the second largest source of greenhouse gas emissions in the U.S. (after transportation). In this homework, you will consider the distributional consequences of imposing a household electricity tax based on the SCC to address the climate change problem.

We recommend using R and writing functions to compute your answers wherever possible. Use the following set of facts:

• Consumers can be separated into two income groups: “high” and “low.” The data set provides
price (in $) and quantity (in kWh) estimates of demand per month for the two groups. Run
linear regressions (with an intercept) to estimate the demand curves for “high” and “low”
income consumers.
• Initially, there is no tax on electricity consumption.
• The current electricity price (without any taxes) is $.10 per kWh.
• The marginal cost of producing a kWh of electricity is linear and has a price-intercept of 0.


# Question 1

One kWh of electricity emits 0.85 pounds of CO2. Assuming that the interim SCC correctly reflects the total social cost of one metric ton of CO2, what is the marginal external cost per kwH of electricity?

To find the MEC we use metric conversion to go from $/ton to cents/kWh:

- (5,100 cents/ 1 metric ton)(1 metric ton/2,205 lbs)(0.85lbs/1kWh)= 1.96 cents/kWh

The marginal external social cost of carbon is **$0.02/kWh**

# Question 2

What is the aggregate monthly demand curve for electricity? What is the supply curve for 
electricity? What is the “benefit” to consumers under the status quo? What is the “benefit” to 
producers under the status quo? What is the environmental cost under the status quo?

```{r}
model_demand_l <- lm(price_cents  ~ q_low_kwh, data=df)
model_demand_h <- lm(price_cents ~ q_high_kwh, data=df)
```

```{r}
summary(model_demand_l)
summary(model_demand_h)
```
Equation of lines:

- Low income demand curve: P = 23.4 - 0.00011Q

- High income demand curve: P= 31.6 - 0.000052Q

- Aggregate demand: P = 31.6 - 0.000052Q [Q<], P= 

```{r}
# need to rearrange the parameter to get Q(P)! 
#Qgg = Qlow(P) + Qlow(h) 

# Importantly, since they-intercepts are different, we know that Qagg(P) will have a kink. I include an ifelse() statement to take
# care of the kink.
```


```{r}
# define a function to get demand

demand <- function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
```


```{r}
# for each p level, return estimated aggregate demand
demand_agg <- function(p){
  q <- demand(p, model_demand_l) + demand(p, model_demand_h)
  return(q)
}
```


```{r}
demand_indv<- function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  return(q)
}
```


```{r}
price_low = seq(0, 23.3, length.out = 100)
demand_low <- demand(price_low, model_demand_l)

low_demand <- map(price_low,demand_low) %>% unlist()

df_2 <- tibble(demand_low = demand_low, price_low = price_low)

ggplot(df_2, aes(demand_low, price_low)) +
  geom_line() + 
  labs(title = "Low Income Supply and Demand for Solar Panels", x = "low income demand", y = "low income price") + 
  theme_light()
```

```{r}
price_high = seq(0, 31.6, length.out = 100)
demand_high <- demand(price_high, model_demand_h)

high_demand <- map(price_high,demand_high) %>% unlist()

df_3 <- tibble(demand_high = demand_high, price_high = price_high)

ggplot(df_2, aes(demand_high, price_high)) +
  geom_line() + 
  labs(title = "title", x = "high demand", y = "high price") + 
  theme_light()
```


```{r}
price = seq(0, 31.6, length.out = 100)
Qagg <- map(price, demand_agg) %>% unlist()

df<- tibble(Qagg = Qagg, price = price)

ggplot(df, aes(Qagg, price)) +
  geom_line() + 
  labs(title = "title", x = "Aggregate Demand Curve", y = "Price") + 
  theme_light()

```

```{r}
# Calculate the supply curve for electricity. We know the current price is $0.10/kWh, and the MC electricity curve is linear and will pass through the 0 intercept. 
# Therefore, the slope of the supply curve can be found by doing (y2-y1/x2-x2) using the points (0,0) and (demand_agg(10),10). The zeros cancel out so we get:

mpc_slope <- (10/(demand_agg(10)))
mpc_slope

supply_p <- function(q){
  p <- mpc_slope*q
  return(q)
}

supply_q <- function(p){
  q <- (p/mpc_slope)
  return(q)
}
```
```{r}
demand_agg(10)
```


```{r}
supply_price = seq(0, 31.6, length.out = 100)
q_supply <- supply_q(price_high)

supply_something<- map(price_high,q_supply) %>% unlist()

df_4 <- tibble(q_supply = q_supply, supply_price = supply_price)

ggplot(df_4, aes(q_supply,supply_price)) +
  geom_line() + 
  labs(title = "title", x = "Q Supply", y = "Price of Supply") + 
  theme_light()

```


```{r}
ggplot() +
  geom_line(df, mapping=aes(Qagg, price, color = "blue"))+
  annotate("text", x = 400000, y = 17,
           label = "Aggregate Demand", angle = -30)+
  geom_line(df_2, mapping=aes(demand_low,price_low, color = "purple")) +
  annotate("text", x = 300000, y = 15,
           label = "High Income Demand", angle = -38)+
  geom_line(df_3, mapping=aes(demand_high,price_high, color = "red")) +
  annotate("text", x = 150000, y = 10,
           label = "Low Income Demand", angle = -55)+
  geom_line(df_4, mapping=aes(q_supply,supply_price, color = "yellow"))+
  annotate("text", x = 300000, y = 7,
           label = "Supply", angle = 18)+
  labs(x = "Quantity of Electricty (kWh)",
       y = "Price per kWh",
       title = "Monthly Demand Curve for Electricity") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(legend.position = "none")
```


```{r}
#Find the q free market value at the market equilibrium point 
#We know that the free market is at equilibrium where p=10, because the current cost of electricity is 10 cents/kwh
Q_f <- demand_agg(10)
P_f <- 10
```

```{r}
# I also define functions for calculating the consumer surplus:

CS <- function(p, model){
  q <- demand(p, model)
  cs <- 0.5*(model$coefficients[[1]] - p)*q
  return(cs)
}

CS_agg <- function(p){
  cs <- CS(p,model_demand_l) + CS(p,model_demand_h)
  return(cs)
}
```


```{r}
#calculate benefit to consumers under status quo"

consumer_surplus_low <- CS(10,model_demand_l)
consumer_surplus_high <- CS(10,model_demand_h)

consumer_surplus_low
consumer_surplus_high

consumer_surplus_agg <- CS_agg(10)
consumer_surplus_agg

```

```{r}
#function for MPC
MPC<- function(q){
  p <- mpc_slope*q
  return(p)
}

PS <- function(q){
  ps <- 0.5*(MPC(q))*q
  return(ps)
}

```

```{r}
supply_q(10)
PS(536719.5)
producer_surplus_f <- PS(supply_q(10))
producer_surplus_f

```

```{r}
#environmental cost 
environmental_cost <- Q_f*1.96
environmental_cost
```
# Question 3

How is the current consumer benefit divided between “high” and “low” 
income consumers?

```{r}
# Find the CS for each income group

cs_l <- CS(10, model_demand_l)
cs_h <- CS(10, model_demand_h)
```

High income consumers have higher benefit when compared to low income consumers. 

- The consumer benefit for low income consumers is $`r round(cs_l/100, 2)`.

- The consumer benefit for high income consumers is $`r round(cs_h/100, 2)`.


# Question 4
4. Derive the optimal electricity tax (in cents per kWh) using the interim SCC. Noting that recent research has shown the poor face a disproportionate share of the impacts from climate change, assume that the climate externality is borne entirely by the “low” income group. What would be the effects of this tax on:
(a) The amount of electricity produced and consumed
(b) The price of electricity
(c) Overall welfare of “high” income consumers
(d) Overall welfare of “low” income consumers
(e) Power suppliers (i.e., electricity producers)
(f) Total environmental damage
(g) Total tax revenue generated

```{r}
#First we need to find the marginal social cost including the externality 

#the optimal tax is equal to the MEC. Therefore the Marginal social cost is just 1.96 cents

#Shift the aggregate demand curve down to account for tax. 


demand_p_t <-function(q,model){
  p <- (model$coefficients[[1]])- 1.96 + (model$coefficients[[2]]*q)+
  return(p)
}

```

```{r}
#demand_new_low <- function(p, model){
 # q <- (p - (model$coefficients[[1]]-t))/model$coefficients[[2]]
  #q <- ifelse(q<0,0,q)
  #return(q)
#}   

demand_new<- function(p, model){
  q <- (p - (model$coefficients[[1]]-1.97))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}   

demand_agg_new <- function(p){
  q <- demand_new(p, model_demand_l) + demand_new(p, model_demand_h)
  return(q)
}

demand_agg_new_eq <- function(q)(demand_new_low(p, model_demand_l) + demand(p, model_demand_h))

```

```{r}
#creating the data frame for the new 
price_new = seq(0, 29.5, length.out = 100)
Qagg_new <- map(price_new, demand_agg_new) %>% unlist()

df_new <- tibble(Qagg_new = Qagg_new, price_new = price_new)
```

```{r}
ggplot()+
  geom_line(df_new, mapping=aes(Qagg_new,price_new, color = "yellow"))
```

**A:** The amount of electricity produced and consumed (increases or decreases) by r round(... , 0) kWh. The new total consumption is r round(...)kWh.

```{r}
#map the new demand curve and the old demand curve to make sure I did
ggplot() +
  geom_line(df, mapping=aes(Qagg, price, color = "Old Agg Demand"))+
  geom_line(df_new, mapping=aes(Qagg_new, price_new, color = "New Agg demand")) +
  scale_color_manual(values = c('red', 'blue')) +
  ylim(0,50) + 
  xlim(0,1000000)+
  labs(x = "Quantity of Electricty (kWh)",
       y = "Price per kWh",
       title = "Monthly Demand Curve for Electricity") + 
  theme_minimal()
```

```{r}
mpc_eq <- function(q)(mpc_slope*q)
```


```{r}
# First I need to find the new demand curve by shifting the aggregate demand curve down by the optimal tax

#demand-supply
#uniroot(function_you_write, c(0,600000))$root
```

```{r}
# First I need to find the new demand curve by shifting the aggregate demand curve down by the optimal tax
equilibrium_pt_free_mkt <- function(p)((demand(p, model_demand_l) + demand(p, model_demand_h))-(p/mpc_slope))
p_free_market <- uniroot(equilibrium_pt_free_mkt, c(0,600000))$root
q_free_market <- demand_agg(p_free_market)
q_free_market
```
```{r}
#calculating the intercept points
equilibrium_pt_tax_51 <- function(p)((demand_new(p, model_demand_l) + demand_new(p, model_demand_h))-(p/mpc_slope))
p_tax_51 <- uniroot(equilibrium_pt_tax_51, c(0,600000))$root
p_tax_51

q_tax_51 <- demand_agg_new(p_tax_51)
q_tax_51

```

**B:** The price of electricity (increases or decreases) by $r round(..., 2). 
Therefore the new price is $r p_new/kWh.

```{r}
# Writing a new function for  consumer surplus:

CS_new <- function(p, model){
  q <- demand_new(p, model)
  cs <- 0.5*(model$coefficients[[1]] - p)*q
  return(cs)
}

CS_agg <- function(p){
  cs <- CS_new(p,model_demand_l) + CS_new(p,model_demand_h)
  return(cs)
}
```


```{r}
#calculate the welfare effects to low and high income consumers 
consumer_surplus_low_t_51 <- CS((p_tax_51+1.96),model_demand_l)
consumer_surplus_high_t_51 <- CS((p_tax_51+1.96),model_demand_h)

consumer_surplus_low_t_51
consumer_surplus_high_t_51

#however, low income consumers will bear the brunt of the environmental cost, so we need to calculate the total enviromental cost.

MEC <- 1.96

environmental_cost_t_51 <- MEC*q_tax_51
environmental_cost_t_51

consumer_surplus_low_true <- consumer_surplus_low-environmental_cost_t_51
consumer_surplus_low_true

```

**C:** The overall welfare of "high" income consumers (with cost of taxes included) decreases by $`r round(consumer_surplus_high_t_51, 0)`.

**D:** The overall welfare of "low" income consumers (with cost of taxes included) decreases by $`r round(consumer_surplus_low_t_51, 0)`. 
*Side note: If we assume that low-income consumers bear the full cost of environmental damage associated with electricity production, then welfare was already negative before the tax was implemented.*

```{r}
PS <- function(q){
  ps <- 0.5*(MPC(q))*q
  return(ps)
}
# calculate the producer surplus 
ps_change <- PS(q_tax_51)
```

**E:** Power suppliers will have a producer surplus decrease by $`r round(ps_change, 0)`.

```{r}
#calculate environmental damage --> I did this already 
environmental_cost_t_51
```

**F:** The total environmental damage will decrease by $`r round(environmental_cost_t_51, 0)`.

```{r}
#calculate tax revenue 
tax_revenue_t_51 <- q_tax_51*1.96
tax_revenue_t_51
```

**G:** The total tax revenue generated will be $`r round(tax_revenue_t_51, 0)`. The tax revenue generated is equivalent to the total environmental cost.

# Question 5 
5. Now, assume that all revenue from the electricity tax will be redistributed to the consumers in proportion to their pre-tax consumption. For example, if 80% of the electricity was consumed by “high” income consumers, then they get 80% of the tax revenue. Additionally, consider the fact that current scientific evidence suggests the true SCC may be much higher than 51. For a range of SCC values (51, 75, 100, 125, and 150 per metric ton of CO2), calculate the effects of an SCC-based electricity tax on:
(a) Overall welfare of “high” income consumers
(b) Overall welfare of “low” income consumers
(c) Electricity producers


```{r}
# Calculate electricity consumption before tax

low_consume <- demand(model_demand_l, p = 10)/((demand(model_demand_h, p = 10)) + demand(model_demand_l, p=10))
# 0.2260845 = 22.61%

high_consume <- demand(model_demand_h, p = 10)/((demand(model_demand_h, p = 10)) + demand(model_demand_l, p=10))
# 0.7739155 = 77.39%

h_welfare <- (consumer_surplus_high_t_51 + high_consume * tax_revenue_t_51)/100
l_welfare <- (consumer_surplus_low_t_51 + low_consume * tax_revenue_t_51)/100
ps_question5 <- ps_change/100
```

```{r}
# Create new data frame with the different price/kWh for different SCC values (I couldn't figure out how to do this, so I did it by hand below. I just know that we should, can someone help me please)

# Calculate the different price/kWh for different SCC values.
# $75/ton SCC value = $75/(2205lbs/$0.85) = $0.0289/kWh
# $100/ton SCC value = $100/(2205lbs/$0.85) = $0.0385/kWh
# $125/ton SCC value = $125/(2205lbs/$0.85) = $0.0482/kWh
# $150/ton SCC value = $150/(2205lbs/$0.85) = $0.0578/kWh

# price_scc_75 <- 2.89
# price_scc_100 <- 3.85
# price_scc_125 <- 4.82
# price_scc_150 <- 5.78

```

####  Social Cost of Carbon at $51/ton

**A:** The overall welfare to “high” income consumers is $`r round(h_welfare, 2)`.

**B:** The overall welfare of “low” income consumers is $`r round(l_welfare, 2)`. This indicates that for low income consumers there is a net loss.

**C:** The overall welfare to electricity producers is $`r round(ps_question5/100, 3)`


#### Social Cost of Carbon at $75/ton

```{r}
# to calculate the "high" income consumers welfare under the tax rebate, it is CS + tax rebate 

# I think these are the CS but I am struggling to find tax rebate. I know this code is not complete yet, I am just trying to figure things out. 
price_CW_75 <- 2.89 
price_CW_100 <- 3.85  
price_CW_125 <- 4.82  
price_CW_150 <- 5.78 
```

```{r}
#write a new demand function where the social cost of carbon is $75 dollars a ton
demand_new_75<- function(p, model){
  q <- (p - (model$coefficients[[1]]-(price_CW_75)))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}   

demand_agg_new_75 <- function(p){
  q <- demand_new_75(p, model_demand_l) + demand_new(p, model_demand_h)
  return(q)
}

equilibrium_pt_tax_75 <- function(p)((demand_new_75(p, model_demand_l) + demand_new_75(p, model_demand_h))-(p/mpc_slope))
p_tax_75 <- uniroot(equilibrium_pt_tax_75, c(0,700000))$root
p_tax_75

q_tax_75 <- demand_agg_new_75(p_tax_51)
q_tax_75

consumer_surplus_low_t_75 <- CS((p_tax_75+price_CW_75),model_demand_l)
consumer_surplus_high_t_75 <- CS((p_tax_75+price_CW_75),model_demand_h)

#calculate the tax revenue 
tax_revenue_t_75 <- q_tax_75*(price_CW_75)

#calculate producer surplus
ps_t_75 <- PS(q_tax_75)

#use tax revnue and consumer surplus to 
h_welfare_75 <- (consumer_surplus_high_t_75 + high_consume * tax_revenue_t_75)/100
l_welfare_75 <- (consumer_surplus_low_t_75 + low_consume * tax_revenue_t_75)/100

```

```{r}
#write a new demand function where the social cost of carbon is $100 dollars a ton
demand_new_100<- function(p, model){
  q <- (p - (model$coefficients[[1]]-(price_CW_100)))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}   

demand_agg_new_100 <- function(p){
  q <- demand_new_100(p, model_demand_l) + demand_new_100(p, model_demand_h)
  return(q)
}

equilibrium_pt_tax_100 <- function(p)((demand_new_100(p, model_demand_l) + demand_new_100(p, model_demand_h))-(p/mpc_slope))
p_tax_100 <- uniroot(equilibrium_pt_tax_100, c(0,600000))$root
p_tax_100

q_tax_100 <- demand_agg_new_100(p_tax_100)
q_tax_100

consumer_surplus_low_t_100 <- CS((p_tax_100+price_CW_100),model_demand_l)
consumer_surplus_high_t_100 <- CS((p_tax_100+price_CW_100),model_demand_h)

#calculate the tax revenue 
tax_revenue_t_100 <- q_tax_100*(price_CW_100)

#calculate producer surplus
ps_t_100 <- PS(q_tax_100)

#use tax revnue and consumer surplus to 
h_welfare_100 <- (consumer_surplus_high_t_100 + high_consume * tax_revenue_t_100)/100
l_welfare_100 <- (consumer_surplus_low_t_100 + low_consume * tax_revenue_t_100)/100
```
```{r}
#write a new demand function where the social cost of carbon is $125 dollars a ton
demand_new_125 <- function(p, model){
  q <- (p - (model$coefficients[[1]]-(price_CW_125)))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}   

demand_agg_new_125 <- function(p){
  q <- demand_new_125(p, model_demand_l) + demand_new_125(p, model_demand_h)
  return(q)
}

equilibrium_pt_tax_125 <- function(p)((demand_new_125(p, model_demand_l) + demand_new_125(p, model_demand_h))-(p/mpc_slope))
p_tax_125 <- uniroot(equilibrium_pt_tax_125, c(0,600000))$root
p_tax_125

q_tax_125 <- demand_agg_new_125(p_tax_125)
q_tax_125

consumer_surplus_low_t_125 <- CS((p_tax_125+price_CW_125),model_demand_l)
consumer_surplus_high_t_125 <- CS((p_tax_125+price_CW_125),model_demand_h)

#calculate the tax revenue 
tax_revenue_t_125 <- q_tax_125*(price_CW_125)

#calculate producer surplus
ps_t_125 <- PS(q_tax_125)

#use tax revnue and consumer surplus to 
h_welfare_125 <- (consumer_surplus_high_t_125 + high_consume * tax_revenue_t_125)/100
l_welfare_125 <- (consumer_surplus_low_t_125 + low_consume * tax_revenue_t_125)/100
```

```{r}
#write a new demand function where the social cost of carbon is $150 dollars a ton
demand_new_150<- function(p, model){
  q <- (p - (model$coefficients[[1]]-(price_CW_150)))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}   

demand_agg_new_150 <- function(p){
  q <- demand_new_150(p, model_demand_l) + demand_new_150(p, model_demand_h)
  return(q)
}

equilibrium_pt_tax_150 <- function(p)((demand_new_150(p, model_demand_l) + demand_new_150(p, model_demand_h))-(p/mpc_slope))
p_tax_150 <- uniroot(equilibrium_pt_tax_150, c(0,700000))$root
p_tax_150

q_tax_150 <- demand_agg_new_150(p_tax_150)
q_tax_150

consumer_surplus_low_t_150 <- CS((p_tax_150+price_CW_150),model_demand_l)
consumer_surplus_high_t_150 <- CS((p_tax_150+price_CW_150),model_demand_h)

#calculate the tax revenue 
tax_revenue_t_150 <- q_tax_150*(price_CW_150)

#calculate producer surplus
ps_t_150 <- PS(q_tax_150)

#use tax revnue and consumer surplus to 
h_welfare_150 <- (consumer_surplus_high_t_150 + high_consume * tax_revenue_t_150)/100
l_welfare_150 <- (consumer_surplus_low_t_150 + low_consume * tax_revenue_t_150)/100
```


# Question 6 
6. Suppose the “high” income group has access to expensive home solar generation. This lowers the electricity demand curve for the “high” income group by half (vertically). Under this new
demand:

(a) What is total electricity consumption?

```{r}
demand_new_solar<- function(p, model){
  q <- (p - (.5*model$coefficients[[1]]))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}   

demand_agg_new_solar <- function(p){
  q <- demand(p, model_demand_l) + demand_new_solar(p, model_demand_h)
  return(q)
}

equilibrium_pt_tax_150 <- function(p)((demand(p, model_demand_l) + demand_new_solar(p, model_demand_h))-(p/mpc_slope))
p_solar <- uniroot(equilibrium_pt_tax_150, c(0,700000))$root
p_solar

q_solar <- demand_agg_new_solar(p_tax_150)
q_solar

```

(b) What is the total environmental externality?
```{r}
environmental_cost_solar<- MEC*q_tax_solar
environmental_cost_solar
```


(c) What value of the electricity tax makes the total environmental damage the same as the
damage when solar panels are available to the high income group?

- The total environmental damage the area under the MEC evaluated at Q. So, we know that the Q value

```{r}

```



# Abi notes
Stuff we still have to do:
Adjust graphs in question 2 to not look bad, and add labels to the curves -- Hollie updated the labels but did not label them cause idk what they're labeled. 
Fill in the right code portions for A and B 
Do questions 5 and 6
Remove the parts of the code that show up in the html that don't need to be there (this should be done at the end)
